---
title: "Territories"
output: html_document
date: "`r Sys.Date()`"
---

# Can we identify (core) territories of individual Wild Boar in their resting/sleeping sites and feeding grounds?

We expect that we can at least to some extent identify core territories. Given that, we further expect that these territories differ in size between the resting sites (the forest patch) and the feeding grounds (agricultural sites).


```{r, warning=F, message=F}
rm(list = ls())
```

```{r, warning=F, message=F}
library(ComputationalMovementAnalysisData)
library(tidyverse)
library(sf)
library(tmap)
library(terra)

```

```{r, warning=F, message=F}
ws <- wildschwein_BE
metadata <- wildschwein_metadata
```

```{r}
names(ws)
names(metadata)
n_distinct(ws$TierID)
n_distinct(ws$CollarID)
```
# Data Preparation
```{r}
ws <- merge(ws, metadata[, c("TierID","Sex")], by = "TierID")
ws <- merge(ws, metadata[, c("TierID","Study_area")], by = "TierID")
  
ws_fin <- ws %>% 
  subset(select = - c(CollarID, day, moonilumination)) %>% 
  filter(Study_area == "Bern")
```

```{r}
# remove ws whole data set (to save storage)
rm(ws)

```



# Transforming to SF
```{r}
ws_sf <- st_as_sf(ws_fin,
                coords = c("E", "N"), 
                crs = 2056)
```

```{r}
ws_sf$TierID <- as.factor(ws_sf$TierID)
ws_grp <- ws_sf %>% group_by(TierID)
ws_smry <- summarise(ws_grp)

mcp <- st_convex_hull(ws_smry)

```




# Plotting Convex Hull's for All 19 Individuals
```{r, message=F, warning=F}
pk100_BE <- terra::rast("pk100_BE.tif")
pk100_BE
```

```{r}
tm_shape(pk100_BE) + 
    tm_rgb() +
    tm_shape(mcp) +
    tm_polygons("TierID", alpha = 0.5, border.col = "red")

```



# Plotting Convex Hull's of Males and females separately
```{r}
# ws_grp -> is a grouped sf object containing all individuals
# m_grp -> just males
# f_grp -> just females

m_grp <- ws_grp %>% filter(Sex == "m")
f_grp <- ws_grp %>% filter(Sex == "f")

m_smry <- summarise(m_grp)
f_smry <- summarise(f_grp)

m_mcp <- st_convex_hull(m_smry)
f_mcp <- st_convex_hull(f_smry)
```


```{r}
#males
tm_shape(pk100_BE) + 
    tm_rgb() +
    tm_shape(m_mcp) +
    tm_polygons("TierID", alpha = 0.5, border.col = "red")
```


```{r}
#females
tm_shape(pk100_BE) + 
    tm_rgb() +
    tm_shape(f_mcp) +
    tm_polygons("TierID", alpha = 0.5, border.col = "red")
```


# Plotting Convex Hull's of Males, minus 48 & 1
```{r}
m_grp <- ws_grp %>% filter(Sex == "m") %>% filter(TierID %in% c(25,31,40,97))

m_smry <- summarise(m_grp)

m_mcp <- st_convex_hull(m_smry)

tm_shape(pk100_BE) + 
    tm_rgb() +
    tm_shape(m_mcp) +
    tm_polygons("TierID", alpha = 0.5, border.col = "red")
```


# Trying Raster Heatmap 










