---
title: "Trajectories_WildBoarData"
output: html_document
date: "`r Sys.Date()`"
---

```{r, warning=F, message=F}
rm(list = ls())
```


```{r, warning=F, message=F}
library(ComputationalMovementAnalysisData)
library(tidyverse)
library(sf)
library(readr)        
library(dplyr)        
library(ggplot2)      
library(lubridate)
library(ggpubr)
```


```{r, warning=F, message=F}
ws <- wildschwein_BE
metadata <- wildschwein_metadata
metadata
```


```{r}
ws <- ws %>% 
  group_by(TierID) %>% 
  mutate(timelag = as.integer(difftime(lead(DatetimeUTC), DatetimeUTC, units = "secs"))) %>% 
  ungroup()

summary(ws$timelag)
# 19 NA's because of 19 distinct ID's
```

```{r}

ws$TierID <- as.factor(ws$TierID)

# Filtering and Removing some columns:
ws_range <- ws %>% 
  filter(timelag >= 0 & timelag < 30000) %>% 
  subset(select = - c(day, CollarID, moonilumination))
```

```{r}
ws$TierID <- as.factor(ws$TierID)

p1 <- ws %>% ggplot(aes(x = DatetimeUTC, y = TierID, colour = TierID)) +
  geom_point(show.legend = F)

p1 
```


```{r}
# 60 second sampling interval
ws_60s <- ws_range %>% filter(timelag >= 50 & timelag < 70)
```

```{r}
# During the relevant and overlapping time
ws_60s <- ws_60s %>% filter(DatetimeUTC > "2015-01-01" & DatetimeUTC < "2016-01-01")


# Plot with the 5 overlapping individuals
p3 <- ws_60s %>% ggplot(aes(x = DatetimeUTC, y = TierID, colour = TierID)) +
  geom_point(show.legend = F)

p3

```


```{r}
range(ws$timelag, na.rm = T)
nrow(ws)

ws_range <- ws %>% filter(timelag >= 0 & timelag < 30000) 

range(ws_range$timelag, na.rm = T)
nrow(ws_range)

nrow(ws) - nrow(ws_range)
```


```{r}
ws_range %>% 
  ggplot(aes(x = timelag)) +
  geom_histogram(binwidth = 60) +
  scale_y_log10()
```
```{r}
w_traj <- st_as_sf((ws_60s), coords = c("E", "N"), crs = 2056, remove = FALSE)
class(w_traj)
```


```{r}
derivates <- function(ws_60s){
    # timelag
    sf_data <- mutate(ws_60s, timelag = as.integer(difftime(lead(ws_60s$DatetimeUTC), ws_60s$DatetimeUTC, units = "secs")))
    # steplength
    ws_60s$steplength <- sqrt(((ws_60s$E - lead(ws_60s$E))**2) + ((ws_60s$N - lead(ws_60s$N))**2)) 
    # speed
    ws_60s$speed <- (ws_60s$steplength / ws_60s$timelag)
    
    ws_60s
}
```

```{r}
ws_60resampled<- derivates(ws_60s)
```


```{r}
traj_map <- function(original, resampled, E, N){
    plot <- ggplot() + 
        geom_path( data = original,  mapping = aes({{E}}, {{N}}, colour = "cyan"), size = 0.8, alpha = 0.2) +
        geom_point(data = original,  mapping = aes({{E}}, {{N}}, colour = "cyan"), alpha = 0.2) + 
        geom_path( data = resampled, mapping = aes({{E}}, {{N}}, colour = "red"),  size = 1) +
        geom_point(data = resampled, mapping = aes({{E}}, {{N}}, colour = "red"),  size = 1.2) +
        theme_bw() +
        theme(legend.justification = "right") +
        theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.title = element_text(face = "bold", hjust = 0.5))

    plot
}
```

```{r}
traj <- traj_map(ws_60s,ws_60resampled, E, N) + 
    scale_colour_discrete(name = "Trajectory", labels = c("1 minute", "3 minutes")) +
    labs(title = "1-Minute vs. 3-Minute sampling interval")
traj
```
```{r}
traj_map <- function(original, E, N){
    plot <- ggplot() + 
        geom_path( data = original,  mapping = aes({{E}}, {{N}}, colour = "cyan"), size = 0.8, alpha = 0.2) +
        geom_point(data = original,  mapping = aes({{E}}, {{N}}, colour = "cyan"), alpha = 0.2) + 
        theme_bw() +
        theme(legend.justification = "right") +
        theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.title = element_text(face = "bold", hjust = 0.5))

    plot
}
```

```{r}
traj <- traj_map(ws_60s, E, N) + 
    scale_colour_discrete(name = "Trajectory", labels = c("1 minute")) +
    labs(title = "1-Minute sampling interval")
traj
```
```


```{r}
plot<- ggplot(data = ws_60s, aes(E, N, colour = TierID)) +
        geom_path(size = 0.8, alpha = 0.3) +
        geom_point(aes( alpha = 0.2))
plot
```


```{r}
sex <- merge(ws_60s, metadata[, c("TierID","Sex")], by = "TierID")
ws <- merge(ws_60s, metadata[, c("TierID","Study_area")], by = "TierID")
  
ws_filter <- sex %>% 
  filter(Sex == "f")
```

```{r}
ws_60s %>% 
  mutate(dt_rounded = round_date(DatetimeUTC, unit = "60 seconds")) -> ws_60s 
```

```{r}
ws_60s %>% filter(TierName == "Caroline") -> ws_caroline
ws_60s %>% filter(TierName == "Miriam") -> ws_miriam
ws_60s %>% filter(TierName == "Olga") -> ws_olga
inner_join(ws_caroline, ws_olga, by = "dt_rounded", suffix = c("_caroline", "_olga")) %>% 
      mutate(distance = sqrt((E_caroline - E_olga)^2 + (N_caroline - N_olga)^2),
           meet = distance < 30) -> ws_join
```


```{r}
ggplot() +
  geom_point(data = ws_caroline, aes(x = E, y = N, color = TierName),  alpha = 0.1, size = 0.5) +
  geom_point(data = ws_olga, aes(x = E, y = N, color = TierName), alpha = 0.1, size = 0.5) +
  geom_point(data = ws_join %>%  filter(meet == TRUE),
             aes(x=E_caroline, y=N_caroline, fill = TierName_caroline), pch = 21, color = "black") +
  geom_point(data = ws_join %>%  filter(meet == TRUE),
             aes(x=E_olga, y=N_olga, fill = TierName_olga), pch = 21, color = "black") #+ theme_void()
```

```{r}
inner_join(ws_caroline, ws_miriam, by = "dt_rounded", suffix = c("_caroline", "_miriam")) %>% 
      mutate(distance = sqrt((E_caroline - E_miriam)^2 + (N_caroline - N_miriam)^2),
           meet = distance < 30) -> ws_join
```

```{r}
ggplot() +
  geom_point(data = ws_caroline, aes(x = E, y = N, color = TierName),  alpha = 0.1, size = 0.5) +
  geom_point(data = ws_miriam, aes(x = E, y = N, color = TierName), alpha = 0.1, size = 0.5) +
  geom_point(data = ws_join %>%  filter(meet == TRUE),
             aes(x=E_caroline, y=N_caroline, fill = TierName_caroline), pch = 21, color = "black") +
  geom_point(data = ws_join %>%  filter(meet == TRUE),
             aes(x=E_miriam, y=N_miriam, fill = TierName_miriam), pch = 21, color = "black") #+ theme_void()
```



```{r}
inner_join(ws_olga, ws_miriam, by = "dt_rounded", suffix = c("_olga", "_miriam")) %>% 
      mutate(distance = sqrt((E_olga - E_miriam)^2 + (N_olga - N_miriam)^2),
           meet = distance < 30) -> ws_join
```

```{r}
ggplot() +
  geom_point(data = ws_olga, aes(x = E, y = N, color = TierName),  alpha = 0.1, size = 0.5) +
  geom_point(data = ws_miriam, aes(x = E, y = N, color = TierName), alpha = 0.1, size = 0.5) +
  geom_point(data = ws_join %>%  filter(meet == TRUE),
             aes(x=E_olga, y=N_olga, fill = TierName_olga), pch = 21, color = "black") +
  geom_point(data = ws_join %>%  filter(meet == TRUE),
             aes(x=E_miriam, y=N_miriam, fill = TierName_miriam), pch = 21, color = "black") #+ theme_void()
```


```{r}
sex <- merge(ws_60s, metadata[, c("TierID","Sex")], by = "TierID")
ws <- merge(ws_60s, metadata[, c("TierID","Study_area")], by = "TierID")
  
ws_filter <- sex %>% 
  filter(Sex == "m")
```


```{r}
ws_60s %>% filter(TierID == "40") -> ws_40
ws_60s %>% filter(TierID == "48") -> ws_48
inner_join(ws_40, ws_48, by = "dt_rounded", suffix = c("_40", "_48")) %>% 
      mutate(distance = sqrt((E_40 - E_48)^2 + (N_40 - N_48)^2),
           meet = distance < 500) -> ws_join
```


```{r}
ggplot() +
  geom_point(data = ws_40, aes(x = E, y = N, color = TierName),  alpha = 0.1, size = 0.5) +
  geom_point(data = ws_48, aes(x = E, y = N, color = TierName), alpha = 0.1, size = 0.5) +
  geom_point(data = ws_join %>%  filter(meet == TRUE),
             aes(x=E_40, y=N_40, fill = TierID_40), pch = 21, color = "black") +
  geom_point(data = ws_join %>%  filter(meet == TRUE),
             aes(x=E_48, y=N_48, fill = TierID_48), pch = 21, color = "black") #+ theme_void()
```




```{r}
# 5 minutes sampling interval
ws_5min <- ws_range %>% filter(timelag >= 590 & timelag < 610)
```

```{r}
# During the relevant and overlapping time
ws_5min <- ws_5min %>% filter(DatetimeUTC > "2015-01-01" & DatetimeUTC < "2016-01-01")


# Plot with the 5 overlapping individuals
p5 <- ws_5min %>% ggplot(aes(x = DatetimeUTC, y = TierID, colour = TierID)) +
  geom_point(show.legend = F)

p5

```

```{r}

```


# detect meet ups between the animals
# then, after a meet up, detect if distance between the animals increased again


