---
title: "Trajectories_2"
output: html_document
date: "`r Sys.Date()`"
---

Trajectories_2

```{r, warning=F, message=F}
rm(list = ls())
```


```{r, warning=F, message=F}
library(ComputationalMovementAnalysisData)
library(tidyverse)
library(sf)
library(readr)        
library(dplyr)        
library(ggplot2)      
library(lubridate)
library(ggpubr)
library(data.table)
```


```{r, warning=F, message=F}
ws <- wildschwein_BE
metadata <- wildschwein_metadata
metadata

ws <- merge(ws, metadata[, c("TierID", "Sex")], by = "TierID") 
```


```{r}
ws <- ws %>% filter(TierID == c(10,22,36,40,48)) 

ws <- ws %>% subset(select = - c(day, CollarID, moonilumination))

ws <- ws %>% mutate(dt_rounded = round_date(DatetimeUTC, unit = "minute"))

```


```{r}
ws_caroline <- ws %>% filter(TierID == 10)
ws_miriam <- ws %>% filter(TierID == 22)
ws_olga <- ws %>% filter(TierID == 36)
ws_franz <- ws %>% filter(TierID == 40)
ws_amos <- ws %>% filter(TierID == 48)

ws_caroline_olga_join <- inner_join(ws_caroline, ws_olga, by = "dt_rounded", suffix = c("_caroline", "_olga")) %>% 
                 mutate(distance = sqrt((E_caroline - E_olga)^2 + (N_caroline - N_olga)^2), meet = distance < 30)

# ws_franz_amos_join <- inner_join(ws_franz, ws_amos, by = "dt_rounded", suffix = c("_franz", "_amos")) %>% 
#                  mutate(distance = sqrt((E_franz - E_amos)^2 + (N_franz - N_amos)^2), meet = distance < 30)

```


```{r}
ggplot() +
  geom_point(data = ws_caroline, aes(x = E, y = N, color = TierName),  alpha = 0.1, size = 0.5) +
  geom_point(data = ws_olga, aes(x = E, y = N, color = TierName), alpha = 0.1, size = 0.5) +
  geom_point(data = ws_caroline_olga_join %>%  filter(meet == TRUE), aes(x = E_caroline, y = N_caroline, fill = TierName_caroline), pch = 21, color = "black") +
  geom_point(data = ws_caroline_olga_join %>%  filter(meet == TRUE), aes(x = E_olga, y = N_olga, fill = TierName_olga), pch = 21, color = "black") #+ theme_void()

```

# -> that's the result without the filter to points with a 60 second interval! -> Caroline was monitored for a long time with mutliple different sampling intervals.

```{r, warning=F, message=F}
rm(list = ls())
```

```{r, warning=F, message=F}
ws <- wildschwein_BE
metadata <- wildschwein_metadata
```


```{r, warning=F, message=F}
# difftime 
ws <- ws %>% group_by(TierID) %>% mutate(timelag = as.integer(difftime(lead(DatetimeUTC), DatetimeUTC, units = "secs"))) %>% 
  ungroup()

# filter to 60 seconds interval
ws <- ws %>% filter(timelag >= 50 & timelag < 70)

ws <- ws %>% subset(select = - c(day, CollarID, moonilumination))

ws <- ws %>% mutate(dt_rounded = round_date(DatetimeUTC, unit = "minute"))

```




# with filter down to 60 second sampling interval:

```{r}
# ws <- ws %>% filter(TierID == c(10,22,36,40,48))
# c(10,22,36,40,48) #those with 60s sampling interval

ws_caroline <- ws %>% filter(TierID == 10)
ws_miriam <- ws %>% filter(TierID == 22)
ws_olga <- ws %>% filter(TierID == 36)
ws_franz <- ws %>% filter(TierID == 40)
ws_amos <- ws %>% filter(TierID == 48)

ws_caroline_olga_join <- inner_join(ws_caroline, ws_olga, by = "dt_rounded", suffix = c("_caroline", "_olga")) %>% 
                 mutate(distance = sqrt((E_caroline - E_olga)^2 + (N_caroline - N_olga)^2), meet = distance < 30)

ws_franz_amos_join <- inner_join(ws_franz, ws_amos, by = "dt_rounded", suffix = c("_franz", "_amos")) %>% 
                 mutate(distance = sqrt((E_franz - E_amos)^2 + (N_franz - N_amos)^2), meet = distance < 30)

```


```{r}
ggplot() +
  geom_point(data = ws_caroline, aes(x = E, y = N, color = TierName),  alpha = 0.1, size = 0.5) +
  geom_point(data = ws_olga, aes(x = E, y = N, color = TierName), alpha = 0.1, size = 0.5) +
  geom_point(data = ws_caroline_olga_join %>%  filter(meet == TRUE), aes(x = E_caroline, y = N_caroline, fill = TierName_caroline), pch = 21, color = "black") +
  geom_point(data = ws_caroline_olga_join %>%  filter(meet == TRUE), aes(x = E_olga, y = N_olga, fill = TierName_olga), pch = 21, color = "black") #+ theme_void()

```

```{r}
meets_caroline_olga <- ws_caroline_olga_join %>% filter(meet == TRUE)

```

# Now code plot with geom_path for each of these meetup groups for both - use meets_caroline_olga for that ?


```{r}
ws_caroline_olga_join <- ws_caroline_olga_join %>% mutate(
  #
  mMinus6 = shift(meet, 6, type = "lead"),
  mMinus5 = shift(meet, 5, type = "lead"),
  mMinus4 = shift(meet, 4, type = "lead"),
  mMinus3 = shift(meet, 3, type = "lead"),
  mMinus2 = shift(meet, 2, type = "lead"),
  mMinus1 = shift(meet, 1, type = "lead"),
  mPlus1  = shift(meet, 1, type = "lag"),
  mPlus2  = shift(meet, 2, type = "lag"),
  mPlus3  = shift(meet, 3, type = "lag"),
  mPlus4  = shift(meet, 4, type = "lag"),
  mPlus5  = shift(meet, 5, type = "lag"),
  mPlus6  = shift(meet, 6, type = "lag"),
  #
)


```


```{r}
meet_traj <- ws_caroline_olga_join %>% filter( mMinus6 + mMinus5 + mMinus4 + mMinus3 + mMinus2 + mMinus1 + 
                                                 mPlus1 + mPlus2 + mPlus3 + mPlus4 + mPlus5 + mPlus6  >  0)

# setting session timezone to UTC (doesn't work otherwise)
Sys.setenv(TZ = "UTC")

occasion1 <- meet_traj %>% filter(dt_rounded > "2015-09-20 05:00:00 UTC" & dt_rounded < "2015-09-20 06:00:00 UTC")
occasion2 <- meet_traj %>% filter(dt_rounded > "2015-09-20 18:00:00 UTC" & dt_rounded < "2015-09-20 21:00:00 UTC")
occasion3 <- meet_traj %>% filter(dt_rounded > "2015-09-20 19:00:00 UTC" & dt_rounded < "2015-09-20 20:00:00 UTC")

# resetting session timezone
Sys.setenv(TZ = "")
```


# Plotting the trajectories

```{r}
ggplot(data = occasion1) +
  geom_path(aes(x = E_caroline, y = N_caroline, fill = TierName_caroline), color = "red") +
  geom_path(aes(x = E_olga, y = N_olga, fill = TierName_olga), color = "cyan") +
  geom_point(aes(x = E_caroline, y = N_caroline, fill = TierName_caroline), pch = 21, color = "black") +
  geom_point(aes(x = E_olga, y = N_olga, fill = TierName_olga), pch = 21, color = "black") +
  geom_text(aes(x = E_caroline, y = N_caroline, label = rownames(occasion1)), nudge_x = -0.5, nudge_y = 4, check_overlap = TRUE, color = "darkred") +
  geom_text(aes(x = E_olga, y = N_olga, label = rownames(occasion1)), nudge_x = -0.5, nudge_y = -4, check_overlap = TRUE, color = "darkblue")
  
  


```

# Occasion 2 (Maybe add 30 or sth to each E_caroline, to make it a bit more visible)
```{r}
ggplot(data = occasion2) +
  geom_path(aes(x = E_caroline, y = N_caroline, fill = TierName_caroline), color = "red") +
  geom_path(aes(x = E_olga, y = N_olga, fill = TierName_olga), color = "cyan") +
  geom_point(aes(x = E_caroline, y = N_caroline, fill = TierName_caroline), pch = 21, color = "black") +
  geom_point(aes(x = E_olga, y = N_olga, fill = TierName_olga), pch = 21, color = "black") +
  geom_text(aes(x = E_caroline, y = N_caroline, label = rownames(occasion2)), nudge_x = 2, nudge_y = 2, check_overlap = TRUE, color = "darkred") +
  geom_text(aes(x = E_olga, y = N_olga, label = rownames(occasion2)), nudge_x = -5, nudge_y = 2, check_overlap = TRUE, color = "darkblue")

```


# Occasion 3
```{r}
ggplot(data = occasion3) +
  geom_path(aes(x = E_caroline, y = N_caroline, fill = TierName_caroline), color = "red") +
  geom_path(aes(x = E_olga, y = N_olga, fill = TierName_olga), color = "cyan") +
  geom_point(aes(x = E_caroline, y = N_caroline, fill = TierName_caroline), pch = 21, color = "black") +
  geom_point(aes(x = E_olga, y = N_olga, fill = TierName_olga), pch = 21, color = "black") +
  geom_text(aes(x = E_caroline, y = N_caroline, label = rownames(occasion3)), nudge_x = 0.5, nudge_y = 3, check_overlap = TRUE, color = "darkred") +
  geom_text(aes(x = E_olga, y = N_olga, label = rownames(occasion3)), nudge_x = -2, nudge_y = 2, check_overlap = TRUE, color = "darkblue")


```


-> we see around the middle numbers where they are closest to each other. 
However, as these two females seem to spend some minitues in close proximity, it looks like there is either no aggression or these two animals know and tolerate each other. ?? We assume that these females are not in the same family, as we saw in the territorial analysis, that they seem to have a different territory.














